Untwisting the Tetrahelix (v0.10)

### 2017-03-14 doc/TetrahelixUntwist.tex
adding links
proofread and clarifications

\title{Untwisting the Tetrahelix (v0.9)}
\author{Robert L. Read \texttt{read.robert@gmail.com} \and
  Robert Gatliff \texttt{robert@toubat.org}
}

@@

\begin{abstract}
  The Boerdijk--Coxeter helix (BC helix, or tetrahelix) is a
  face-to-face stack of regular tetrahedra.  Considering the edges of
  these tetrahedra as structural members, the resulting structure is attractive and
  inherently rigid, and therefore interesting to architects,
  mechanical engineers, and robotocists.  A formula that matches the
  visually apparent helices forming the outer rails of the tetrahelix
  is derived which is convenient for designers.  This formula defines
  the vertices of tetrahelices of varying radius, pitch, and
  curvature, with the BC helix as a special case.  An addtional
  special case is of that of $0$ curvature or rail angle, which
  generates a \emph{tetrabeam}.  A particular choice of paramaters
  defines a novel object, the minimax optimal member length-difference
  tetrabeam, the \emph{equitetrabeam}.  A formula for computing the
  radius and z-axis travel that gives an optimal minimax tetrahelix
  for any rail angle between $0$ and $\rho_{bc}$ is given, thus
  defining a continuum of tetrahelices terminated by the equitetrabeam
  and BC helix. Numerically finding the rail angle from the equation for
  pitch allows optimal tetrahelices of any pitch to be designed. 
  An interactive tool for such design and experimentation is provided.
  A formula
  for the inradius is given.
  Utility and use for static and variable geometry
  truss/space frame design and robotics based on the are discussed.
\end{abstract}

@@

The Boerdijk--Coxeter helix\cite{coxeter1985simplicial} (BC helix), is
a face-to-face stack of tetrahedra that winds about a straight axis.
Because architects, structural engineers, and robotocists are inspired
by and follow such mathematical models but can build structures and
machines of differing or even dynamically changing length, it is
useful to develop the mathematics of structure formed from tetrahedra
where we relax regularity.

The vertices of the tetrahedra lie upon
three helices about the central axis.  The
Tetrobot/Glussbot\cite{TetrobotBook} project uses the regularity of
this geometry to make a tentacle-like robot that can crawl like a slug
or mollusc.  The Tetrobot concept is to use mechanical members, called
actuators, which can change their length, connected by special joints,
such as the three-d printable Song-Kwon-Kim\cite{song2003spherical}, or the CMS joint\cite{HamlinSandersonCMS}
which allow many members to come to a single point.
Such machines can
follow purely regular mathematical models such as the Boerdijk–-Coxeter
helix or the Octet Truss\cite{richard1961synergetic}.

Buckminster Fuller called the BC helix a \emph{tetrahelix}\cite{fuller1982synergetics},
a term now commonly used. In this paper we reserve \emph{BC helix} to mean the purely regular structre and use \emph{tetrahelix} to refer
to any structure isomophic to a the BC helix, whether regular or not.

@@

Imagining Figure \ref{fig:closeup} as a static mechanical structure,
we observe that it is useful to the mechanical engineer or
robotocist because the structure remains an inherently rigid,
omni-triangulated space frame, which may be expected to be at least
somewhat mechanically strong.
Imagine further in Figure \ref{fig:closeup}, that each static edge was replaced with an
actuator that could dynamically become shorter or longer in response to electronic control,
and the they vertices were a joint that supported sufficient angular displacement
for this to be possible. Such a machine is a Tetrobot\cite{TetrobotBook}.

BC helix does not rest on a plane in a simple way. It is convenient to
be able to ``untwist'' it and to form a tetrahelix space frame that has a
flat planar surface. By making length changes in a certain way, we can
untwist a tetrahelix to form a \emph{tetrabeam} which has planar faces
and has, for example, an equilateral triangular profile. This paper
develops the equations needed to untwist the tetrahelix. All math
deveoped here is available in JavaScript and demonstrated by an interactive
design website \url{https://pubinv.github.io/tetrahelix/}\cite{readtetrahelix},
from which Figures \ref{fig:closeup} and
the figure below are taken.

In Figure \ref{fig:spudboy}, the closest helix is the BC helix, and the furthest
is the equitetrabeam.

\section{A User's Formulation of the BC Helix}

If you can choose member lengths, you can form a linear combination of
the equitetrabream lengths and the completely regular lengths of the
tetrahelix, thereby choosing the amount of twist\footnote{The formal definiton
  of twistiness, or \emph{torsion}, is not useful or used in this paper, and
  we use \emph{curvature} without relying on the mathematical curvature easily
  computed from helical equations we provide.}.
If you are designing a
space frame, this is a static design choice, in a robot, it is a
dynamic choice that can be used to twist the robot and/or exert linear or
angular force on the environment.

Ideally we would have a simple formula for defining the nodes based on
any curvature or pitch we choose.  Unfortunately, it is not obvious that a linear
combination of lengths produces a simple formula.  It is a goal of
this paper to relate these two approaches to generating a tetrahelix
continuum.

@@

R.W. Gray's site\cite{graytetrahelix}, repeating a formula by Coxeter\cite{coxeter1985simplicial} in more accessible form, gives for a counter-clockwise BC Helix:
\begin{equation}
  \label{graycoxeter}
V(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos(n \theta) $\\
   $ r_{bc} \sin(n \theta) $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}
where $n$ represents each integer numbered node in succession on every colored rail.

@@

Such a helix can be written:
\begin{equation}
  \label{eq:colord}
H_{BCcolored}(n,c) =
\left [
  \begin{tabular}{c}
   $ r_{bc}  \cos((3 \theta - 2 \pi)n + c  \theta $\\
   $ r_{bc} \sin((3 \theta - 2 \pi)n + c  \theta $\\
   $ 3 h_{bc} (n + c/3)  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}

In this formula, integral values of $n$ may be taken as a node number for one rail and used to compute its Cartesian
coordinates. Allowing $n$ to take non-integer values defines a continuous
helix in space which is close to the segmented polyline of the outer tetrahedra edges, and coincides with them at integer
values.

@@

Note in Figure \ref{railanglefig} the $z$-axis travel for one rail edge is denoted by $d$. In \eqref{graycoxeter} and \eqref{eq:colored}, the variable
$h$ is used for one third of the distance we name $d$. We will later justify that $d = 3h$. In this paper we assume the length of a rail
is alwyas $1$ is a simplification, although we make the rail length a parameter in our JavaScript codein \texttt{tetrahelixmath.js} \cite{readtetrahelix}.

This Gray-Coxeter formulation $V(n)$ from \eqref{graycoxeter} is valuable, but obscures the fact that we have three
apparent rail helices.
In order to rewrite this expression with an explicit rotation of $2\pi/3$, we expand 
the expressions $\theta$ and $\rho_{bc}$ and seek to isolate the term $c2\pi/3 $.
\begin{align*}
  \rho_{bc} n + c \theta  &=   \text{\{we aim for 3 in denominator, so we split...\}} \\
    (3 \theta - 2 \pi)n + (c/3)  (3 \theta)  &=   \text{\{we want $2\pi$ in numerator, so add canceling terms...\}} \\
  (3 \theta - 2 \pi)n + (c/ 3) (3 \theta - 2 \pi  + 2 \pi) &=  \text{\{associate...\}} \\
  (3 \theta - 2 \pi)n + (c/ 3) ((3 \theta - 2 \pi)  + 2 \pi) &=  \text{\{distribute...\}} \\  
  (3 \theta - 2 \pi)n + (c / 3) (3 \theta - 2 \pi)  + c 2 \pi /3 &=  \text{\{definition of $\rho_{bc}$...\}} \\
  \rho_{bc} n + (c / 3) \rho_{bc}  + c 2 \pi /3 &=  \text{\{collect like factors...\}} \\  
  n \rho_{bc}  +\rho_{bc} ( c+  2 \pi) /3  \\
\end{align*}
Now the the term on the left is the only one that depends on the scalar $n$. We use this to a create
a new formulation: $H_{BCsymmetric}(n,c) = H_{BCcolored}(n,c)$

@@

%% Actually, I am not sure if this is a clearer formulation...it might be, and need to put it in general!!!
\begin{equation}
H_{BCsymmmetric}(n,c) =
\left [
  \begin{tabular}{c}
    $ r  \cos(\rho_{bc} n + c (\rho_{bc} +  2 \pi) /3) $\\
   $ r  \sin(\rho_{bc} n + c (\rho_{bc} +  2 \pi) /3) $\\
   $ (n+c/3)  h_{bc} $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
%%  $\kappa = n + c/3$ \\
    $\rho_{bc} = (3 \theta - 2 \pi)$ \\
    $ \theta = \arccos(-2/3) $ \\
    $ h_{bc} = 1/\sqrt{10} $ \\    
  \end{tabular}      
\end{equation}

Since:
\[ \frac{2 \pi}{\rho_{bc}} \approx 10.16
\]
We can see that there are approximately $10.16$ red, blue or yellow tetrahedra on one rail in a single revolution.
The pitch of the Boerdijk--Coxeter helix of edge length $1$ is the length of three tetrahedra times this number:
\begin{align*}
  &= \frac{d_{bc} 2 \pi }{\rho_{bc}} \\  
%%  &= \frac{3 \cdot h_{bc} 2 \pi }{\rho_{bc}} \\
  &= \frac{3  \sqrt{\frac{2}{5}}  \pi}{\rho_{bc}} \\
  &\approx 9.64 \\
\end{align*}
The pitch is less than the number of tetrahedra because the tetrahedra
are not lined up perfectly.  It is a famous and interesting result
that the pitch is irrational. A BC helix never has two tetrahedra at
precisely the same orientation around the $z$-axis. However, this is
inconvenient to designers, who might prefer a rational pitch.
The idea of developing a rational period by arranging solid tetrahedra by relaxing the face-to-face matching
has been explored\cite{sadler2013periodic}. In this paper
we develop slightly irregular edge lengths that support, for example, a pitch of precisely 12
tetrahedra in one revolution which would allow an architect to design a
column having a basis and a capital in the same relation to the
tetrahedra they touch at the bottom and top of the column.

@@

Since the axes are parallel, we may define the \emph{inradius} of a
tetrahelix to be the radius of the largest
cylinder parallel to this axis contained within the circumscribing cylinder w
hich is penetrated by no edge.
Define an \emph{minimax edge-length optimal tetrahelix} or just an
\emph{optimal tetrahelix} to be a tetrahelix for which there exists
no other tetrahelix of the same inradius and pitch with a lower maximum
difference between its edge lengths.

@@

Since all three rails have the same rail length, no matter how we
move the rails in the $xy$ plane if we shorten the $xy$ distance between
vertices we shorten the total distance.
Our tool for thinking about this is to project out the $z$ dimension to form
a two-dimensional figure of nodes and non-rail edges.
Consider the projection along the $z$ axis of all vertices and non-rail edges into the $xy$ plane, which will be
a figure of dots and connecting segments in the $xy$ plane. The convex
hull for any one helix projection will be a circle (if its pitch is
irrational) or a polygon if rational, or a point if the helix has
zero curvature. Each of these figures by definitions lies outside the
circle of inradius $i$ in the $xy$ plane.

@@

Case 1: Suppose that $\rho$ is zero. Then for any given inradius, an equilateral
triangle is the minimax solution for all non-rail edges. Since there is only
one rail-edge length, this is the minimax solution for the entire set. Since
the projection is an equilateral triangle lies on a circle, all points lie on
a cylinder.

@@

Supoose there is a rail $P$ which does not have a segment touching the incircle
at all in the projection, as depicted in 
Then a segment connecting the other two $Q$ and $R$ is a longest
rail, because any chord touching an incircle is longer than any which does not.
These two rails ($Q$ and $R$) can be moved closer to each other, and further away from $P$, to form a better minimax solution
by shortening the longest rail, until one of the segments from $P$ the previously
untouching rail touches one of the incircle. By induction, then all rails have
at least one segement touching each vertex on the rail that is tangent to the
incircle in an optimal tetrahelix.

@@

    Any attempt to move any rail in either $z$ direction lengthens one two-hop edge to $t' > t$
    edge and shortens one one-hop edge $o' < o$. Let $B = \{o',t' \} \cup A$ be new edges.
    The minimax of $B$ > $A$ since there is a single rail length which cannot be both greater
    than $t'$ and $o'$ and less than $t'$ and $o'$.
    Therefore, any optimal tetrahelix has all one-hop edges between all rails are equal to each other, and
    all two-hop edges are equal to each other, and the $z$ distances between rails are equal, and therefore
    $d/3$ from each other.
\end{proof}

 Note that based on \ref{eventhirds}, there are only 3 possible lengths in an optimal tethrahelix,
 and we are justified in classifying edge lengths as \emph{rail},\emph{one-hop}, or
\emph{two-hop}. The one-hop edges are the edges between closest on the $z$-axis, and the two-hop edges are those that hop over a vertex.

@@

$H_{general}$ forces the user to select an $d_{\rho}$
which has a sensible radius.

@@

By Theoerm \ref{eventhirds}, we can compute the (at most) three edge-lengths of an optimal
tetrahelix by (where $dist$ is the cartesian distance function):
\begin{align*}
  \text{rail} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n+1,c,\rho,d_{\rho}),r_{\rho})) = 1 \\
  \text{one-hop} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n,c+1,\rho,d_{\rho},r_{\rho}))  \\
  \text{two-hop} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n,c+2,\rho,d_{\rho},r_{\rho}))  \\  
\end{align*}
Which are invarinat for all $n$ and $c$.

@@

If you look down the axis of an optimal tetrahelix, it happens that only the one-hop edges
come closest to the axis. In other words, they define the radius of the incircle of the
projection, or the radius of a cylinder that would just fit inside the tetrahelix.
This is useful to know if the tetrahelix is used as a structure to house a pipe or
a ladder that must bear a human or to armor a conduit. The inradius $i_\rho$ of
an optimal tetrahelix is a remarkably simple function of the radius $r$ and the rail angle $\rho$:
\begin{equation}
  \label{eq:inradius}
  i_{\rho} = \frac{r \sin{(\pi - \rho)}}{6}
\end{equation}
Which can be seen from the trigonometry of a diagram of the projected one-hop edges
connecting four sequentially numbered vertices:

@@

\section{The Equitetrabeam}

@@

This radius\footnote{Another interesting but non-optimal solution is derived by setting (one-hop + two-hop)/2 = 1,
  occurs at $r = \sqrt{35}/4$ which produces
  three length classes of $11/12, 12/12, 13/12$.}
produces a two-hop rail length of $\frac{2}{\sqrt{3}}$.
The difference between this and $1$ is $\approx 15.47\% $.

@@

We observe that Equations \eqref{eqrhoopt} and \eqref{dopt} compute $r_{opt}$ and $d_{opt}$ which
create an optimal tetrahelix for any value rail angle $\rho$ between $0$, which
gives the Equitetrahelix and
$\rho_{bc} \approx 35.43 \degree$, which gives the BC Helix.

 Because the equitetrabeam which has a rail angle of $0$ still has
 chirality, that is, one still must decide to connect with one-hop to
 the clockwise or counter-clockwise node, it is not poosible to build
 a smooth continuum where $\rho$ transitions from positive to negative
 which remains optimal. One can use a negative $\rho$ in $H_{general}$
 but it does not produce minimax optimal tetrahelices. In other words,
 untwisting a conter-clockwise tetrahelix to rail angle $0$ and then going
even further does produce a clockwise tetrahelix, but one in which the
 one-hop and two-hop lengths in the wrong places (that is, two-hop
 becomes shorter than one-hop.) Likewise, $\rho > \rho_{bc}$ generates
 a tetrahelix, but minimax optimality is not guaranteed by $H_{general}$.

The pitch of any tetrahelix 
where $\rho \neq 0$ is:
\begin{equation}
  \label{pitcheqn}
p(\rho) = \frac{2 \pi  \cdot d}{\rho}
\end{equation}
For a fixed $z$-axis travel $d$, this is trivial.
However, if one is computing 
$z$-axis travel from \eqref{dopt} the pitch is not simple.
However, it increases monotonically and smoothly with decreasing $\rho$, so
Equation \eqref{pitcheqn} can be easily solved numerically with a Newton-Raphson
solver, as we do on our website.
For a pitch at least $ p \geq \frac{3  \sqrt{\frac{2}{5}}  \pi}{\rho_{bc}} \approx 9.64 $,
using \eqref{dopt} produces minimax optimal tetrahelices.

@@

Perhaps surprisingly, the optimal untwisting is accomplished only by
changing the length of the two-hop member, leaving the one-hop memever
and rail length equivalent within this continuum.\footnote{Before deriving Equation \eqref{eqrhoopt}, we created a continuum by
using a linear interpolation between the optimal radius for the
Equitetrabeam and the BC Helix. This minimax optimum of this simpler
approach was at most 1\% worse than the optimum computed by
\eqref{eqrhoopt}.}
 However, it should
be noted that an engineer or architect may also use $H_{general}$
directly and interactively, and that minimax length optimality is a
mathematic starting point rather than the final word on the beauty and utility of
physical structures. For example, a structural engineer might increase
radius in order to resist buckling. If an equitetrabeam were actually
used as a beam, an engineer might start with the optimal tetrabeam and
dilate it in one dimension to ``deepen'' the beam. Similarly, simple
changes curve the equitetrabeam into an ``arch''.

@@

>>>>

\title{Untwisting the Tetrahelix (v0.10)}
\author{Robert L. Read \texttt{read.robert@gmail.com} \and
  Robert Gatliff \texttt{robert@toubat.org}
}

@@

\begin{abstract}
  The Boerdijk--Coxeter helix (BC helix, or tetrahelix) is a
  face-to-face stack of regular tetrahedra.  Considering the edges of
  these tetrahedra as structural members, the resulting structure is attractive and
  inherently rigid, and therefore interesting to architects,
  mechanical engineers, and robotocists.  A formula that matches the
  visually apparent helices forming the outer rails of the BC helix.
  This formula is generalized to a formula convenient to designers.
  Formulae for 
  computing the
  parameters that give an edge-length minimax-optimal tetrahelices
  are given, defining a continuum of tetrahelices of varying curvature.
  The endpoints of the optimality of this continuum are the BC helix and
  a structure of zero curvature, the \emph{equitetrabeam}.
  Numerically finding the rail angle from the equation for
  pitch allows optimal tetrahelices of any pitch to be designed. 
  An interactive tool for such design and experimentation is provided: \url{https://pubinv.github.io/tetrahelix/}.
  A formula for the inradius of optimal tetrahelices is given.
  Utility and use for static and variable geometry
  truss/space frame design and robotics based on the are discussed.
\end{abstract}

@@

The Boerdijk--Coxeter helix\cite{coxeter1985simplicial} (BC helix), is
a face-to-face stack of tetrahedra that winds about a straight axis.
Because architects, structural engineers, and robotocists are inspired
by and follow such mathematical models but can build structures and
machines of differing or even dynamically changing length, it is
useful to develop the mathematics of structures formed from tetrahedra
where we relax regularity.

The vertices of the tetrahedra lie upon
three helices about the central axis.  The
Tetrobot/Glussbot\cite{TetrobotBook} project uses the regularity of
this geometry to make a tentacle-like robot that can crawl like a slug
or mollusc.  The Tetrobot concept is to use mechanical members, called
actuators, which can change their length, connected by special joints,
such as the three-d printable Song-Kwon-Kim\cite{song2003spherical} joint, or the CMS joint\cite{HamlinSandersonCMS}
which allow many members to come to a single point.
Such machines can
follow purely regular mathematical models such as the Boerdijk–-Coxeter
helix or the Octet Truss\cite{richard1961synergetic}.

Buckminster Fuller called the BC helix a \emph{tetrahelix}\cite{fuller1982synergetics},
a term now commonly used. In this paper we reserve \emph{BC helix} to mean the purely regular structre and use \emph{tetrahelix} to refer
to any structure isomorphic to a the BC helix.

@@

Imagining Figure \ref{fig:closeup} as a static mechanical structure,
we observe that it is useful to the mechanical engineer or
robotocist because the structure remains an inherently rigid,
omni-triangulated space frame, which is
somewhat mechanically strong.
Imagine further in Figure \ref{fig:closeup}, that each static edge was replaced with an
actuator that could dynamically become shorter or longer in response to electronic control,
and the they vertices were a joint that supported sufficient angular displacement
for this to be possible. Such a machine is a Tetrobot\cite{TetrobotBook}.

A BC helix does not rest comfortably on a plane. It is convenient to
be able to ``untwist'' it and to form a tetrahelix space frame that has a
flat planar surface. By making length changes in a certain way, we can
untwist a tetrahelix to form a \emph{tetrabeam} which has planar faces
and has, for example, an equilateral triangular profile. This paper
develops the equations needed to untwist the tetrahelix. All math
deveoped here is available in JavaScript and demonstrated by an interactive
design website \url{https://pubinv.github.io/tetrahelix/}\cite{readtetrahelix},
from which Figures \ref{fig:closeup} and
the figure below are taken.

In Figure \ref{fig:spudboy}, the closest helix is the BC helix, and the furthest
is the equitetrabeam, defined in Section \ref{sec:equitetrabeam}.

\section{A Designer's Formulation of the BC Helix}

We would like to design nearly regular tetrahelices with a formula that
gives the vertices in space. Eventually we would like to design nearly regular
tetrahelices by choosing the lengths of a small set of members.
If you are designing a
space frame, this is a static design choice, in a robot, it is a
dynamic choice that can be used to twist\footnote{The formal definition of twistieness, or \emph{torsion},
  is not useful or used in the paper. The formal \emph{curvature} and \emph{torsion} of the helices defined here
may be easily computed from the formualae if desired.} the robot and/or exert linear or
angular force on the environment.

Ideally we would have a simple formula for defining the nodes based on
any curvature or pitch we choose. It is a goal of
this paper to relate these two approaches to generating a tetrahelix
continuum.

@@

R.W. Gray's site\cite{graytetrahelix}, repeating a formula by Coxeter\cite{coxeter1985simplicial} in more accessible form, gives for a counter-clockwise BC Helix:
\begin{equation}
  \label{graycoxeter}
V(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos{n \theta} $\\
   $ r_{bc} \sin{n \theta} $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}
where $n$ represents each integer numbered node in succession on every colored rail.

@@

Such a helix can be written:
\begin{equation}
  \label{eq:colored}
H_{BCcolored}(n,c) =
\left [
  \begin{tabular}{c}
   $ r_{bc}  \cos{((3 \theta - 2 \pi)n + c  \theta)} $\\
   $ r_{bc} \sin{((3 \theta - 2 \pi)n + c  \theta)} $\\
   $ 3 h_{bc} (n + c/3)  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}

In this formula, integral values of $n$ may be taken as a node number for one rail and used to compute its Cartesian
coordinates. Allowing $n$ to take non-integer values defines a continuous
helix in space which is close to the segmented polyline of the outer tetrahedra edges, and equals them at integer
values.

@@

Note in Figure \ref{railanglefig} the $z$-axis travel for one rail edge is denoted by $d$. In \eqref{graycoxeter} and \eqref{eq:colored}, the variable
$h$ is used for one third of the distance we name $d$. We will later justify that $d = 3h$. In this paper we assume the length of a rail
is always $1$ is a simplification, although we make the rail length a parameter in our JavaScript code in \texttt{tetrahelixmath.js} \cite{readtetrahelix}.

This Gray-Coxeter formulation $V(n)$ from \eqref{graycoxeter} is valuable, but obscures the fact that we have three
apparent rail helices.
In order to rewrite this expression with an explicit rotation of $2\pi/3$, we expand 
the expressions $\theta$ and $\rho_{bc}$ in \eqref{eq:colored} and seek to isolate the term $c2\pi/3 $.
\begin{align*}
 c \theta  &=   \text{\{we aim for 3 in denominator, so we split...\}} \\
    (c/3)  (3 \theta)  &=   \text{\{we want $2\pi$ in numerator, so add canceling terms...\}} \\
 (c/ 3) ((3 \theta - 2 \pi)  + 2 \pi) &= \text{\{definition of $\rho_{bc}$\}...} \\  
  (c / 3) \rho_{bc}  + c 2 \pi /3 &=  \text{algebra...} \\  
c  ( \rho_{bc} +  2 \pi) /3  \\
\end{align*}
We use this to a create
a new formulation: $H_{BCsymmetric}(n,c) = H_{BCcolored}(n,c)$.
%% TODO: I seem to have lost the thread of my thought. I don't seem to be introducing d at that right time.

@@

%% Actually, I am not sure if this is a clearer formulation...it might be, and need to put it in general!!!
\begin{equation}
H_{BCsymmmetric}(n,c) =
\left [
  \begin{tabular}{c}
    $ r  \cos{\rho_{bc} n + c (\rho_{bc} +  2 \pi) /3} $\\
   $ r  \sin{\rho_{bc} n + c (\rho_{bc} +  2 \pi) /3} $\\
   $ (n+c/3)  h_{bc} $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
%%  $\kappa = n + c/3$ \\
    $\rho_{bc} = (3 \theta - 2 \pi)$ \\
    $ \theta = \arccos(-2/3) $ \\
    $ h_{bc} = 1/\sqrt{10} $ \\    
  \end{tabular}      
\end{equation}

We can see the each rail tetrahedron moves us $\rho_{bc}$ radians around a circle. Since:
\[ \frac{2 \pi}{\rho_{bc}} \approx 10.16
\]
we can see that there are approximately $10.16$ red, blue or yellow tetrahedra on one rail in a single revolution.

The pitch of the Boerdijk--Coxeter helix of edge length $1$ is the length of three tetrahedra times this number:
\begin{align*}
   \frac{3 h_{bc}\cdot  2 \pi }{\rho_{bc}} 
  &= \frac{6 \pi}{\sqrt{10}\rho_{bc}} 
  &\approx 9.64 
\end{align*}

The pitch is less than the number of tetrahedra because the tetrahedra
are not lined up perfectly.  It is a famous and interesting result
that the pitch is irrational. A BC helix never has two tetrahedra at
precisely the same orientation around the $z$-axis. However, this is
inconvenient to designers, who might prefer a rational pitch.
The idea of developing a rational period by arranging solid tetrahedra by relaxing the face-to-face matching
has been explored\cite{sadler2013periodic}. 
We develop below slightly irregular edge lengths that support, for example, a pitch of precisely 12
tetrahedra in one revolution which would allow an architect to design a
column having a basis and a capital in the same relation to the
tetrahedra they touch at the bottom and top of the column.

@@

Since the axes are parallel, we may define the \emph{inradius} of a
tetrahelix to be the radius of the largest
cylinder parallel to this axis contained within the circumscribing cylinder
which is penetrated by no edge.
Define a \emph{minimax edge-length optimal tetrahelix} or just an
\emph{optimal tetrahelix} to be a tetrahelix for which there exists
no other tetrahelix of the same inradius and pitch with a lower maximum
difference between its edge lengths.

@@

Since all three rails have the same rail length, no matter how we
move the rails in the $xy$ plane if we shorten the $xy$ distance between
vertices we shorten the total distance.
Our tool for thinking about this is to project out the $z$ dimension to form
a two-dimensional figure of nodes and non-rail edges (see Figure \ref{untouchingrailfig}.
Consider the projection along the $z$ axis of all vertices and non-rail edges into the $xy$ plane, which will be
a figure of dots and connecting segments in the $xy$ plane. The convex
hull for any one helix projection will be a circle (if its pitch is
irrational) or a polygon if rational, or a point if the helix has
zero curvature. Each of these figures by definitions lies outside the
circle of inradius $i$ in the $xy$ plane.

@@

Case 1: Suppose that $\rho$ is zero. Then for any given inradius, an equilateral
triangle is the minimax solution for all non-rail edges. Since there is only
one rail-edge length, this is the minimax solution for the entire set. Since
the vertices of an equilateral triangle lies on a circle, all points lie on
a cylinder.

@@

Supoose there is a rail $P$ which does not have a segment touching the incircle
at all in the projection, as depicted in Figure \ref{untouchingrailfig}.
Then a segment connecting the other two $Q$ and $R$ is a longest
rail, because any chord touching an incircle is longer than any which does not.
These two rails ($Q$ and $R$) can be moved closer to each other, and further away from $P$, to form a better minimax solution
by shortening the longest rail, until one of the segments from $P$ the previously
untouching rail touches one of the incircle. By induction, then all rails have
at least one segement touching each vertex on the rail that is tangent to the
incircle in an optimal tetrahelix.

@@

    Any attempt to move any rail in either $z$ direction lengthens one two-hop edge to $t'$, where $t' > t$
    edge and shortens one one-hop edge $o' < o$. Let $B = \{o',t' \} \cup A$ be new edges.
    The minimax of $B$ is greater than the minimax of $A$ since there is a single rail length which cannot be both greater
    than $t'$ and $o'$ and less than $t'$ and $o'$.
    Therefore, any optimal tetrahelix has all one-hop edges between all rails are equal to each other, and
    all two-hop edges are equal to each other, and the $z$ distances between rails are equal, and therefore
    $d/3$ from each other.
\end{proof}

 Note that based on Theorem \ref{eventhirds}, there are only 3 possible lengths in an optimal tethrahelix,
 and we are justified in classifying edge lengths as \emph{rail}, \emph{one-hop}, or
\emph{two-hop}. The one-hop edges are the edges between closest on the $z$-axis, and the two-hop edges are those that skip over a vertex.

@@

$H_{general}$ forces the user to select a $d_{\rho}$
which has a sensible radius.

@@

By Theoerm \ref{eventhirds}, we can compute the (at most) three edge-lengths of an optimal
tetrahelix by formula universally quantified by $n$ and $c$:
\begin{align*}
  \text{rail} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n+1,c,\rho,d_{\rho}),r_{\rho})) = 1 \\
  \text{one-hop} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n,c+1,\rho,d_{\rho},r_{\rho}))  \\
  \text{two-hop} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n,c+2,\rho,d_{\rho},r_{\rho}))  \\  
\end{align*}
where $dist$ is the Cartesian distance function.

@@

If you look down the axis of an optimal tetrahelix, it happens that only one of the one-hop edges
come closest to the axis. In other words, they define the radius of the incircle of the
projection, or the radius of a cylinder that would just fit inside the tetrahelix.
This is useful to know if the tetrahelix is used as a structure to house a pipe or
a ladder that must bear a human or to armor a conduit. The inradius $i_\rho$ of
an optimal tetrahelix is a remarkably simple function of the radius $r$ and the rail angle $\rho$:
\begin{equation}
  \label{eq:inradius}
  i_{\rho} = r \sin{\frac{\pi - \rho}{6}}
\end{equation}
Which can be seen from the trigonometry of a diagram of the projected one-hop edges
connecting four sequentially numbered vertices:

@@

From this equation with the help of symbolic computation we observe that inradius of the BC helix of unit rail length is $i_{\rho_{bc}} = \frac{3}{10\sqrt{2}} \approx 0.21213$.

\section{The Equitetrabeam}
\label{sec:equitetrabeam}

@@

This radius\footnote{Another interesting but non-optimal solution is derived by setting
  (one-hop + two-hop)/2 = 1,
  occurs at $r = \sqrt{35}/4$ which produces
  three length classes of $11/12, 12/12, 13/12$.}
produces a two-hop rail length of $\frac{2}{\sqrt{3}}$.
The difference between this and $1$ is $\approx 15.47\% $. The inradius if the equitetrabeam of unit
rail length from both Equation \eqref{eq:inradius} and the fact that the inradius of
an equilateral triangle is half the circumradius is $\sqrt{\frac{8}{27}}/2$, or $\frac{\sqrt{6}}{9}$.

@@

We observe that Equations \eqref{eqrhoopt} and \eqref{dopt} compute $r_{opt}$ and $d_{opt}$ which
create an optimal tetrahelix for any value rail angle $\rho$ between $0$, which
gives the equitetrabeam and
$\rho_{bc} \approx 35.43 \degree$, which gives the BC helix.

 Because the equitetrabeam which has a rail angle of $0$ still has
 chirality, that is, one still must decide to connect the one-hop edge to
 the clockwise or the counter-clockwise node, it is not poosible to build
 a smooth continuum where $\rho$ transitions from positive to negative
 which remains optimal. One can use a negative $\rho$ in $H_{general}$
 but it does not produce minimax optimal tetrahelices. In other words,
 untwisting a conter-clockwise tetrahelix to rail angle $0$ and then going
even further does produce a clockwise tetrahelix, but one in which the
 one-hop and two-hop lengths in the wrong places (that is, two-hop
 becomes shorter than one-hop.) Likewise, $\rho > \rho_{bc}$ generates
 a tetrahelix, but minimax optimality is not guaranteed by $H_{general}$.

The pitch of any tetrahelix 
where $\rho \neq 0$ is:
\begin{equation}
  \label{pitcheqn}
p(\rho) = \frac{2 \pi  \cdot d}{\rho}
\end{equation}
For a fixed $z$-axis travel $d$, this is trivial.
However, if one is computing 
$z$-axis travel from \eqref{dopt} the pitch is not simple.
It increases monotonically and smoothly with decreasing $\rho$, so
Equation \eqref{pitcheqn} can be easily solved numerically with a Newton-Raphson
solver, as we do on our website.
For a pitch at least $ p \geq \frac{3  \sqrt{2}  \pi}{\sqrt{5}\rho_{bc}} \approx 9.64 $,
using \eqref{dopt} produces minimax optimal tetrahelices.

@@

Perhaps surprisingly, the optimal untwisting is accomplished only by
changing the length of the two-hop member, leaving the one-hop member
and rail length equivalent within this continuum.\footnote{Before deriving Equation \eqref{eqrhoopt}, we created a continuum by
using a linear interpolation between the optimal radius for the
Equitetrabeam and the BC Helix. This minimax optimum of this simpler
approach was at most 1\% worse than the optimum computed by
\eqref{eqrhoopt}.}
 However, it should
be noted that an engineer or architect may also use $H_{general}$
directly and interactively, and that minimax length optimality is a
mathematic starting point rather than the final word on the beauty and utility of
physical structures. For example, a structural engineer might increase
radius in order to resist buckling. If an equitetrabeam were actually
used as a beam, an engineer might start with the optimal tetrabeam and
dilate it in one dimension to ``deepen'' the beam. Similarly, simple
changes curve the equitetrabeam into an ``arch''.

@@

Research Turk's knots
